Interval Probability for Fuzzy Quantum Theories

### 2016-05-11 prob.tex
prob.tex \section{Quantum Probability Spaces} updated...

There is however another more subtle source of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly
$\ell$. and the needles lengths are not all equal and are not all
absolutely equal to $\ell$. These perturbations make the events
``fuzzy.'' For example, instead of talking about the idealized event
that exactly $M$ needles cross lines which would require the most
expensive needles built to the most precise accuracy, laser precision
for drawing lines on the floor, and the most powerful microscopes to
determine if a needle does cross a line, we might instead talk about
the event that $M-\delta$ needles evidently cross lines and
$M+\delta'$ needles plausibly cross lines for some acceptable
approximations $\delta$ and $\delta'$. This fuzzy notion of events
leads us to consider intervals of probability reflecting the certainty
of the event and its plausibility. In an actual experiment with
$3204$ needles and the ratio $\frac{\ell}{h}=0.6$~\cite{DeMorgan1872}, 
there are $1213$ needles clearly crossed a line and $11$ ones
was close enough to plausibly be considered as crossing the line: we
would express the probability in this case as the interval
$\left[\frac{1213}{3204},\frac{1234}{3204}\right]$ expressing that we are
certain that the event has probability at least $\frac{1213}{3204}$ but it is possible that
it would have probability $\frac{1234}{3204}$.

@@

\begin{definition}[Projection Operators; Orthogonality;
Commutativity~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Swart2013}]
Given a Hilbert space $\Hilb$, a projection operator $P$ is a linear transformation from $\Hilb$
to itself such that $P^2 = P = P^{\dagger}$. Projection operators have the
following properties:\footnote {``Projection'' is sometimes called
``orthogonal projection" or ``self-adjoint projection" to
emphasize $P^{\dagger} = P$~\cite{Maassen2010}.}
\begin{itemize}
\item Projection operators $P_1$ and $P_2$ are orthogonal if $P_1P_2 =
  P_2P_1 = \mathbb{0}$;
\item Projection operators $P_1$ and $P_2$ commute if $P_1P_2 =
  P_2P_1$;
\item If the projections $P_1$ and $P_2$ are orthogonal then $P_1 +
  P_2$ is also a projection;
\item If the projections $P_1$ and $P_2$ commute then $P_1P_2$ is also
  a projection.
\end{itemize}
\end{definition}

\amr{Here it would be good to refer to the notion of ``quantum test''
  and define events as sums of quantum tests. This will automatically
  include everything except the products of commutative projections
  which we will have to explain that they can be expressed as sums of
  orthogonal projections.}

>>>>

There is however another more subtle source of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly
$\ell$. and the needles lengths are not all equal and are not all
absolutely equal to $\ell$. These perturbations make the events
``fuzzy.'' For example, instead of talking about the idealized event
that exactly $M$ needles cross lines which would require the most
expensive needles built to the most precise accuracy, laser precision
for drawing lines on the floor, and the most powerful microscopes to
determine if a needle does cross a line, we might instead talk about
the event that $M-\delta$ needles evidently cross lines and
$M+\delta'$ needles plausibly cross lines for some acceptable
approximations $\delta$ and $\delta'$. This fuzzy notion of events
leads us to consider intervals of probability reflecting the certainty
of the event and its plausibility. In another experiment with
$3204$ needles and the ratio $\frac{\ell}{h}=0.6$~\cite{DeMorgan1872}, 
there are $1213$ needles clearly crossed a line and $11$ ones
was close enough to plausibly be considered as crossing the line: we
would express the probability in this case as the interval
$\left[\frac{1213}{3204},\frac{1224}{3204}\right]$ expressing that we are
certain that the event has probability at least $\frac{1213}{3204}$ but it is possible that
it would have probability $\frac{1224}{3204}$.

@@

\begin{definition}[Projection Operators; Orthogonality
~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Swart2013}]
Given a Hilbert space $\Hilb$, an event, formally called elementary quantum
test~\cite{peres1995quantum}, can be represented as a projection operator
$P:\Hilb\rightarrow\Hilb$ onto a linear subspace~$S$ of $\Hilb$. The set of all
events can be defined recursively as follow:
\footnote {``Projection'' is sometimes called
``orthogonal projection" or ``self-adjoint projection" to
emphasize $P^{\dagger} = P$~\cite{Maassen2010}.}
\begin{itemize}
\item $\mathbb{0}$ is a projection.
\item For any vector $\ket{\psi}\in \Hilb$,
  $\frac{\proj{\psi}}{\ip{\psi}{\psi}}$ is a projection.
\item Projection operators $P_1$ and $P_2$ are orthogonal if $P_1P_2 =
  P_2P_1 = \mathbb{0}$. The sum of two orthogonal projections~$P_1 +
  P_2$ is also a projection. Moreover, if $P_1$ and $P_2$ project onto
  subspaces $S_1$ and $S_2$ respectively, then $P_1 + P_2$ will project onto the
  linear span of $S_1$ and $S_2$, i.e., $\left\{\ket{\psi_1} + \ket{\psi_2}
  \middle| \ket{\psi_1}\in S_1 \textrm{ and } \ket{\psi_2}\in S_2 \right\}$.
\end{itemize}
\end{definition}

Notice that the preceding definition exhaust all possible projection operators.
Because every linear subspace~$S$ has an orthonormal basis~$\{\ket{\psi_j}\}$,
the projection onto $S$ can always expressed as $\sum_j\proj{\psi_j}$.
As operators, we can also multiply projections together.

\begin{definition}[Commutativity
~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Swart2013}]
Projection operators $P_1$ and $P_2$ commute if $P_1P_2 =
  P_2P_1$;
If the projections $P_1$ and $P_2$ commute then $P_1P_2$ is also
  a projection. Moreover, if $P_1$ and $P_2$ project onto
  subspaces $S_1$ and $S_2$ respectively, then $P_1P_2$ will project onto the
  intersection of $S_1$ and $S_2$, i.e., $S_1 \cap S_2$.
\end{definition}
